\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}


\title{Notes on "Topology and Geometry" by Bredon}
%
\author{Kevin Sony}
\date{Spring 2026}

\newcommand{\norm}[1]{\left\lVert#1\right\rVert}

\newcommand{\betabar}{\bar{\beta}}
\newcommand{\kbar}{\bar{k}}
\newcommand{\R}{\mathbb{R}}

\begin{document}
	
	\maketitle
	
	\section{Metric Spaces}
	
	\subsection{Definition of Metric Space}
	A metric space is a set $X$ with a function $d: X \times X \to \R$ called the metric that satisfies the following properties
	\begin{enumerate}
		\item $d(x, y) \geq 0$ for all $x, y \in X$
		\item $d(x, y) = d(y, x)$ for all $x, y \in X$
		\item $d(x, z) \leq d(x, y) + d(y, z)$ for all $x, y, z \in X$
	\end{enumerate}
	The $\epsilon$ ball about a point $x \in X$ is defined to 
	\begin{gather*}
		B_{\epsilon}(x) = \{ p \in X, d(x, p) < \epsilon \}
	\end{gather*}
	
	\subsection{Proposition}
	A function $f : X \to Y$ between metric spaces is continuous if and only if $f^{-1}(U)$ is open in $X$ for each open subset $U$ of $Y$.
	\\ \\
	\noindent \textit{Proof:}
	Suppose that $f$ is continuous and $U$ is an open subset of $Y$. Let $f(x) \in U$. Since $U$ is an open set, then there exists an open ball $B_{\epsilon}(f(x))$ that is completely contained in $U$. Recall that the a function is continuous if for an arbitrary $\epsilon > 0$, there exists a $\delta > 0$ such that $d_X(x_1, x_2) < \delta$ implies $d_Y(f(x_1), f(x_2)) < \epsilon$. Now consider a point $y_1 = f(x_1) \in B_{\epsilon}(f(x))$. Clearly, $d_Y(y, y_1) = d_Y(f(x), f(x_1)) < \epsilon$ and so $d_X(x_1, x_2) < \delta$, and so there exists a $\delta$ ball about $x \in X$, $B_{\delta}(x)$, and clearly $B_{\delta}(x) \subset f^{-1}(U)$, so $f^{-1}(U)$ is open in $X$.
	
	To show the converse, suppose that $f^{-1}(U)$ is open in $X$ for each open subset $U$ of $Y$. Consider any point $y = f(x)$ and an $epsilon > 0$. The corresponding $\epsilon$ ball is $B_{\epsilon}(y)$. From the assumption, we know there is a corresponding open set $f^{-1} (B_{\epsilon}(y))$. Furthermore, we know that $x \in f^{-1} (B_{\epsilon}(y))$ and so we can deduce that there exists an open $\delta$ ball contained entirely within this set, i.e. $B_{\delta}(x) \subset f^{-1} (B_{\epsilon}(y))$. We thus conclude that for the $\epsilon > 0$, there exists a $\delta > 0$ such that for any point $p \in B_{\delta}(x)$, $f(p) \in B_{\epsilon}(y)$, which is the definition of continuity of $f$.
	
	\subsection{Proposition}
	If $d_1$ and $d_2$ are metrics on the same set $X$ which satisfy the condition that for any two points $x, y \in X$ and $\epsilon > 0$, there is a $\delta > 0$ such that
	\begin{gather*}
		d_1(x, y) < \delta \implies d_2(x, y) < \epsilon \\
		d_2(x, y) < \delta \implies d_1(x, y) < \epsilon
	\end{gather*}
	then the metrics define the same open set on $X$.
	\\ \\
	\noindent \textit{Proof:}
	Suppose the given condition holds for two metrics $d_1$ and $d_2$. We want to show that an open set on one of the metric spaces, say $U \subset (X, d_1)$, is also open on the other metric space. We know that since $U$ is open in $(X, d_1)$, there exists an open ball $B_{\epsilon}(u) \subset U$ defined by $d_1$ for some point $u \in U$. This means that for every point $p \in B_{\epsilon}(u)$, $d_1(p, u) < \epsilon$. From the condition, we know that for a given $\epsilon > 0$, there must exist a $\delta > 0$ such that $d_2(p, u) < \delta \implies d_1(p, u) < \epsilon$, i.e. that the open ball $B_{\delta}(u)$ defined by $d_2$ is contained in $U$. This thus holds for every point $u \in U$. Therefore, $U$ is open in $(X, d_2)$.
	
	\subsection*{Problems}
	
	\subsubsection*{Problem 1}
	Consider the set $C[0, 1]$ of continuous real functions on $[0, 1]$. Show that
	\begin{gather*}
		d(f, g) = \int_0^1 |f(x) - g(x)| dx
	\end{gather*}
	defines a metric on $C[0, 1]$. Does this still hold if continuity is weakened to integrability .
	
	We clearly see that $|f(x) - g(x)| \geq 0$ for $f, g \in C[0, 1]$ and $x \in [0, 1]$ and so
	\begin{gather*}
		d(f, g) \geq 0
	\end{gather*}
	Similarly, $|f(x) - g(x)| = |g(x) - f(x)|$ and so
	\begin{gather*}
		d(f, g) = d(g, f)
	\end{gather*}
	Now, consider $f, g, h \in C[0, 1]$. Recall that for a point $x \in [0, 1]$, $|f(x) - g(x)| \leq |f(x) - h(x)| + |h(x) - g(x)|$ and so
	\begin{gather*}
		d(f, g) \leq d(f, h) + d(h, g)
	\end{gather*}
	Thus, $d$ defines a metric on $C[0, 1]$.
	
	Now consider the set of Riemann integrable functions, $R[0, 1]$. Let $f, g, h \in R[0, 1]$. We know that the Lebesgue’s criterion states that a function $f$ is Riemann integrable if and only if the measure of the set of its discontinuities $\mu(D_f)$ is zero.

	Consider $m = f - g$. We know clearly that $D_f$ and $D_g$ have measure zero, and we also know that $D_m \subset D_f \cup D_g$ meaning $\mu(D_m) = 0$. Therefore, $m$ is Riemann integrable and
	\begin{gather*}
		d(f, g) = \int_0^1 |f(x) - g(x)| dx = \int_0^1 |m(x)| dx \geq 0
	\end{gather*}
	It is self evident that for $f, g \in R[0, 1]$, $d(f, g) = d(g, f)$. Finally, via similar construction as in the first case, we know that the new functions $m = f - g$, $n = f - h$ and $p = h - g$ are also Riemann integrable and so $d(f, g) \leq d(f, h) + d(h, g)$.
	
	\textit{Checking this online, it turns out the definition of the metric space given actually defines a pseudometric space. This is because two distinct functions $f, g$ can still result in $d(f, g) = 0$}
	
	\subsubsection*{Problem 2}
	If $X$ is a metric space and $x_0 \in X$, show that $f: X \to \R$ given by $f(x) = d(x, x_0)$ is continuous.
	
	From the triangle inequality, we know that for the points $x, y, x_0 \in \R$, $d(x, x_0) \leq d(x, y) + d(y, x_0)$ and $d(y, x_0) \leq d(y, x) + d(x, x_0)$ and so $d(x, y) \geq |d(y, x_0) - d(x, x_0)|$, i.e. $|f(x) - f(y)| < d(x, y)$. Thus, $f$ is Lipschitz continuous.
	
	\subsubsection*{Problem 3}
	If $A$ is a subset of a metric space $X$, define a function $f : X \to \R$ given by $f(x) = d(x, A) = \inf_{y \in A} d(x, y)$. Show that $f$ is continuous.
	
	From the definition, we have that
	\begin{gather*}
		|f(x) - f(y)| = |d(x, A) - d(y, A)|
	\end{gather*}
	where
	\begin{gather*}
		d(x, A) = \inf_{a_x \in A} d(x, a_x) \\
		d(y, A) = \inf_{a_y \in A} d(y, a_y)
	\end{gather*}
	Notice that $d(x, A) \leq d(x, a)$ for some arbitrary point $a \in A$, and likewise that $d(y, A) \leq d(y, a)$ for some $a \in A$, and so 
	\begin{gather*}
		|d(x, A) - d(y, A)| \leq |d(x, a) - d(y, a)|
	\end{gather*}
	for some arbitrary $a \in A$
	
	The triangle inequality states that $d(m, p) \leq d(m, n) + d(n, p)$, for some arbitrary points $m, n, p \in X$
	\begin{gather*}
		d(m, p) \leq d(m, n) + d(n, p) \implies d(m, p) - d(m, n) \leq d(n, p) \\
		d(m, n) \leq d(m, p) + d(p, n) \implies d(m, n) - d(m, p) \leq d(p, n)
	\end{gather*}
	Therefore, $|d(m, n) - d(m, p)| \leq d(n, p)$ for arbitrary points $m, n, p \in X$. With that in mind, we have that $|d(x, a) - d(y, a)| \leq d(x, y)$. Therefore, $|f(x) - f(y)| \leq d(x, y)$, and so $f$ is Lipschitz continuous.
	
	\section{Topological Spaces}
	Sometimes, we would like to study spaces spaces without regard to metric, and since the things to study involve open sets, we introduce the concept of the topological space.
	
	\subsection{Definition of Topological Space}
	A topological space is a set $X$ together with a collection of subsets of $X$ called open sets such that
	\begin{enumerate}
		\item the intersection of two open sets is open
		\item the union of any collection of open sets is open
		\item the empty set $\emptyset$ and the whole space $X$ is open
	\end{enumerate}
	A set $C$ is called closed if $X - C$ is open.
	
	\subsection{Definition of Continuity on Topological Spaces}
	If $X$ and $Y$ are topological spaces and $f: X \to Y$ is a function, then $f$ is said to be continuous if $f^{-1}(U)$ is open for every open set $U \subset Y$.
	\\ \\
	\noindent
	It follows from this definition and the definition of closed sets that $f$ is continuous if and only if $f^{-1}(C)$ is closed for every closed set $C \subset Y$.
	\\
	\noindent \textit{Proof:}
	Suppose that $f: X \to Y$ is continuous. Clearly then $f^{-1}(U)$ is open for every open set $U \subset Y$. By definition of closed sets, we can generate every closed set by $C = Y - U$. The final step is to show that $f^{-1}(C)$ is closed. To show this, we first must show that $f^{-1}(Y - U) = X - f^{-1}(U)$. To do this, suppose to the contrary that this was not the case, i.e. either there exists an element $x$ both in $f^{-1}(U)$ and $f^{-1}(Y - U)$, or there exists an element not in either one of them. Consider the first scenario. If $x \in f^{-1}(U)$ and $x \in f^{-1}(Y - U)$, then $f(x) \in U$ and $f(x) \in Y - U$, a contradiction. Now consider the second scenario. If $x \notin f^{-1}(U)$ and $x \notin f^{-1}(Y - U)$, then $f(x) \notin U$ and $f(x) \notin Y - U$, which is also a contradiction. Thus, $f^{-1}(Y - U) = X - f^{-1}(U)$, and since $f^{-1}(U)$ is open, $f^{-1}(Y - U) = f^{-1}(C)$ is closed. This completes the proof.
	
	\subsection{Definition of Neighborhoods}
	If $X$ is a topological space and $x \in X$, then a set $N$ is called a neighborhood of $x$ in $X$ if there exists an open set $U \subset N$ such that $x \in N$.
	
	\subsection{Definition of Neighborhood Basis}
	If $X$ is a topological space and $x \in X$, then a collection $B_x$ of subsets of $X$ containing $x$ is called a neighborhood basis at $x$ in $X$ if each neighborhood of $x$ in $X$ contains some element of $B_x$ and each element of $B_x$ is a neighborhood of $x$.
	

\end{document}